Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-12 paper/example.tex
Revise intro and example

The expected behavior of the SW system, and the components in its implementation, are modeled with AGREE contracts. The contracts constrain input and state properties of output for component models. AGREE model checks this assume-guarantee system to hierarchically prove the composite system obeys all contract obligations under all possible finite input streams.

The initial AGREE specification for the components and SW system make no assumptions about the integrity of the inputs and outputs. The cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously. As such, the AGREE specification for the AI route planner is updated to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output; in other words, its output is unconstrained in the AGREE specification allowing it to take on any value and allowing it to send that value at any time.

The contracts for the other components are also updated with the threat analysis information. For example, the specification for the legacy waypoint manager adds assumptions about about its input being \emph{well-formed} since that is no longer known \emph{a priori} as AI route planner is now untrusted.

Two requirements are added to the SW AGREE specification related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated downstream; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request. Here the goal is to make clear that a delayed or unrequested automation response may prevent the UAV from flying its intended mission or to fly the wrong mission.

AGREE examines the composition of the new specifications for the implementation in \figref{fig:example}(a) to determine whether it complies with the added cyber-resilience properties, by way of model checking. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about input to the legacy waypoint manager. The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request. A clear malicious behavior from the untrusted component.

@@

The system implementation is cyber-hardened by automatically transforming the model to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a). A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating. The auto-generated AGREE specification states that only well-formed inputs are passed to the output. The user must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. It raises an alert if ever the specified temporal properties are violated. The AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and, also,  that response must come with the request or in the next step after the request. As with the filter, the user provides the policy, and that policy is based on the existing AGREE specification in the SW system.

>>>>

The expected behavior of the SW system, and the components in its implementation, are modeled with AGREE contract specifications. The contracts constrain input and state properties of output for component models. AGREE model checks this assume-guarantee system to hierarchically prove the composite system obeys all contract obligations under all possible finite input streams.

The initial AGREE contract specifications for the components and the overall system make no assumptions about the integrity of the inputs and outputs. A resulting cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously. As such, the designer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output; in other words, its output is unconstrained in the AGREE specification allowing it to take on any value and allowing it to send that value at any time.

The contract specifications for the other components are also updated with the threat analysis information by adding in new requirements to mitigate the identified cyber-vulnerabilities. For example, a designer adds to the specification for the legacy waypoint manager assumptions about its input being \emph{well-formed} since that is no longer known \emph{a priori} as the AI route planner is now untrusted. The systems designer is also responsible for defining the meaning of well-formed, and in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude with an additional check on IDs in the message for integrity.

Two other requirements are added to the SW AGREE specification by the designer related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} requires all the waypoints sent to the UAV flight controller to be well-formed (it detects if malformed waypoints are propagated downstream); and the second, \emph{Alert if start is not bounded relative to a request}, requires an automation request to correspond with an automation response either in the same step or within one step (it detects if a mission is being started without a request or if the start is delayed more than one step after a request). Here the goal of the requirements is to detect if the untrusted component is trying to prevent the UAV from flying its intended mission or to fly a wrong mission.

AGREE examines the composition of the new specifications for the implementation in \figref{fig:example}(a) to determine whether it complies with the added cyber requirements, by way of model checking. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about input to the legacy waypoint manager. The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request. A clear malicious behavior from the untrusted component.

@@

The system implementation is cyber-hardened by automatically transforming the model to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a). A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating. The auto-generated AGREE specification states that only well-formed inputs are passed to the output. The system designer must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. It raises an alert if ever the specified temporal properties are violated. The AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and, also,  that response must come with the request or in the next step after the request. As with the filter, the system designer provides the policy, and that policy is based on the existing AGREE specification in the SW system.

### 2021-05-12 paper/intro.tex
Revise intro and example

A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to harden the system against potential cyber-vulnerabilities. The behavior of the model is given by contracts on components expressed in the AGREE modeling language. These contracts are modified as a result of the cyber-vulnerability analysis to assume properties on the input and and guarantee properties of the output that reflect a hardened system. In the absence of any transformation to implement the hardening, AGREE fails to prove the system cyber-hardened.

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, hardened system meets its cyber-resiliency requirements.

The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

@@

>>>>

A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to satisfy requirements added to mitigate cyber vulnerabilities. In this work, the behavior of the system is modeled by contracts on components expressed in the AGREE modeling language. These contracts are modified as a result of the cyber-vulnerability analysis with the new requirements for mitigation.  The requirements assume properties on the input and guarantee properties of the output that reflect a hardened system. In the absence of any transformation to implement such requirements, AGREE fails to prove the system meets these new requirements.

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.

The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML which is a verified implementation of ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

@@

All the aforementioned tools are publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4, demonstration videos, etc. \cite{phase2, camkes} ---ADD LINK TO VIDEO.
